---
title: "Sample analysis"
format:
  html: default
  docx: default
---

```{r setup, include=FALSE}
# Load packages:

library(knitr)
library(here)
library(readr)
library(dplyr)
library(forcats)
library(stringr)
library(tidyr)
library(labelled)
# NOTE: {gtsummary} is heavy; install & adapt code at your own convenience
# library(gtsummary) # TODO: Uncomment (& install packages if using {gtsummary})
# library(broom)         # These do not need to be loaded actually, but they are
# library(broom.helpers) #   needed for {gtsummary} to work.


# Environment configuration:

## Set working directory to the project root (reproducibly):
opts_knit$set(root.dir = here())

# Set chunk options:
opts_chunk$set(echo = FALSE) # Omit syntax from output
```

# Introduction

We will analyze the body mass differences in arctic penguins.

# Methods

We load and preprocess the dataset, and then compute its descriptive statistics, after filtering out non-complete cases.

```{r load-data}
## Read dataset:
penguins_data <- read_csv(
  file = "dat/penguins.csv",
  col_types = cols(
    species  = col_factor(),
    island   = col_factor(),
    sex      = col_factor(),
    year     = col_integer(),
    .default = col_double()
  )
)
```

```{r preprocess-data}
## Recode values:
penguins_data <- penguins_data |>
  mutate(sex = sex |> fct_relabel(str_to_sentence))

## Assign labels:
# (see https://allisonhorst.github.io/palmerpenguins/reference/penguins.html
#   for more info):
penguins_data <- penguins_data |> set_variable_labels(
  species           = "Penguin species",
  island            = "Island in Palmer Archipielago",
  bill_length_mm    = "Bill length (mm)",
  bill_depth_mm     = "Bill depth (mm)",
  flipper_length_mm = "Flipper length (mm)",
  body_mass_g       = "Body mass (g)",
  sex               = "Sex",
  year              = "Study year"
)

## Filter out non-complete cases:
penguins_data <- penguins_data |> drop_na()
```

```{r descriptive-stats}
## Simple summary:
descriptive_table <- penguins_data |> summary()
## TODO: Drop if using neater {gtsummary} output

## Create descriptive statistics table:
# descriptive_table <- penguins_data |> tbl_summary()
## TODO: Uncomment if using {gtsummary}
```

Finally, we try to explain weight by sex and species, fitting a linear model to the individual body mass data.

```{r model-fit}
bodymass_fit <- lm(body_mass_g ~ sex + species, data = penguins_data)
## TODO: Add interaction/covariates? Test model assumptions?

bodymass_coef_table <- bodymass_fit |> coef()
## TODO: Drop if using {gtsummary}

# bodymass_coef_table <- bodymass_fit |> tbl_regression()
## TODO: Uncomment if using {gtsummary}
```

# Results

In the following we see the descriptive statistics.

```{r descriptives-table}
descriptive_table
```

<br>

This table shows the model coefficients.

```{r model-output}
bodymass_coef_table
```
